*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
Fixed reference to PUR section.

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: 'Mireille Savoie'
    institute: mta
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    institute: [mta, ubc]
  - name: Julie Nadeau
    institute: mta
  - name: Laurel Genge
    institute: mta
  - name: Sylwia Śliwińska-Wilczewska
    institute: [mta, ug]
  - name: Maximilian Berthold
    institute: mta
  - name: Naaman M. Omar
    institute: mta
  - name: Ondřej Prášil
    institute: [mta, algatech]
    main: false
  - name: Amanda M. Cockshutt
    institute: [mta, sfx]
    main: false
  - name: Douglas A. Campbell
    institute: mta
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Departments of Biology and Chemistry and Biochemistry, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - ubc: 'Department of Biology, University of British Columbia'
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
  - algatech: 'Institute of Microbiology, Center Algatech, Laboratory of Photosynthesis, Novohradska 237, Trebon, CZ 37981, Czech Republic.'
  - ug: 'Institute of Oceanography, University of Gdansk, 46 Piłsudskiego St, Gdynia, Poland'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: FALSE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA Maine, USA. MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120 (CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773) is from Low-Light adapted (LLIV) clade. Cultures were maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected PAR)in the source niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were transferred weekly with 1 in 5 dilutions with Pro99 media [@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial seawater [@bergesEvolutionArtificialSeawater2001].

@@

Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from Clade LLIII (ex. SS120) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. Photosystem I and II were only detected at low [O~2~]. The exclusive assignment of ATP Synthase peptides to Clade IV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

@@

XXXRewrite caption if we change figure versionXXX
```{r NATLProchloroPhotosynthDepthO2, fig.cap = NATLProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "NATLProchloroPhotosynthDepthO2.png"))
```

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)A), consistent with Murphy *et al*. [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct fits for different photoperiods refer to figure \@ref(fig:AllStrainPurFitsPlots)A.

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:BluevsRedPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~.  Under 25 µM O~2~ red light again generated more growth of SS120 per unit cumulative diel PUR, than did blue light, again consistent with lower cost of growth through lower photoinactivation under red light. For distinct fits for different photoperiods refer to figure \@ref(fig:AllStrainPurFitsPlots)B.

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, lower, flatter response to cumulative diel PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)C).

Our data again support enhanced growth under conditions of low cumulative diel PUR and 660 nm (red) spectral bandwidth, consistent with Murphy *et al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue, wavebands.  Interestingly, this protective effect of red light disappears for MIT9313 growing under 2.5 µM O~2~, possibly because photoinactivation is strongly suppressed under this low [O~2~]. For distinct fits for different photoperiod fits refer to figure \@ref(fig:AllStrainPurFitsPlots)C.

@@

\newpage
XXXXEdit text depending upon which ocean protein figures we useXXXXX
Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLIII SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR.  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313.  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:NATLProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~], suggesting limited requirement for protein turnover under low [O~2~].

@@

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day, and thus this strain will not exploit habitats with short photoperiods, typical of winter or light attenuated depths. MED4 is also excluded from the lowest oxygen habitats, equivalent to our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as demonstrated by the growth under our 25 µM O~2~ experiments. Genomic analyses (Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under red light. TARA Oceans
Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018 did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll *a* concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:NATLProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, at depths up to 200 meters with O~2~ of 15 µM.

*Prochlorococcus marinus* SS120, a LLIII clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by @lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:NATLProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones but again cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the clade LLIII tolerance range.

@@

*Prochlorococcus marinus* MIT9313, a LLIV clade representative, shows the potential to inhabit future warmer temperate zones year-round, as it grows under a 4 h photoperiod, expected in winter or at light-attenuated depths. MIT9313 demonstrates an unexpected tolerance to higher light levels and cumulative diel PUR,  but only under low oxygen conditions of 25 µM and 2.5 µM (Figure \@ref(fig:MIT9313PAR)). The ability to utilize higher PAR and higher cumulative diel PUR under 25 µM or 2.5 µM O~2~ enables MIT9313 to grow in OMZ, even at depths closer to the surface. MIT9313 carries a gene encoding (S)-2-hydroxy-acid oxidase [@cunaneCrystalStructureAnalysis2005],  with a K~M~ for [O~2~] of ~ 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), which produces H~2~O~2~.  Growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffers photoinhibition, resulting from the inactivation of PSII caused by the production of H~2~O~2~. This photoinhibition is compounded by the limited inducible repair for PSII, due to the absence of FtsH 2 and 3  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  MIT9313 shows remarkable ability to thrive under very low [O~2~], potentially allowing it to expand into broader ecological niches. These results are supported by Figure \@ref(fig:NATLProchloroPhotosynthDepthO2) showing PSI protein subunits annotated as derived from MIT9313 detected at depths \> 120 m, along with PSII subunits at depths from 50 m to 200 m in regions where O~2~ was 15 µM.

In future work we aim to test whether MIT9313 is not “using” photosynthesis in low O~2~ environments to drive CO~2~ fixation, so much as to generate oxygen locally to support  respiration. The deep water environments typical for MIT 9313 are relatively nutrient rich and  *Prochlorococcus* take up and metabolise various sugars [@gomez-baenaGlucoseUptakeIts2008; @munoz-marinGlucoseUptakeProchlorococcus2017; @munoz-marinProchlorococcusCanUse2013] and amino acids [@zubkovDepthRelatedAmino2004]. They may be using photosynthesis to generate O~2~ to act as an electron sink for respiration, to generate using ATP for maintenance and to take up nutrients from the surroundings.
Partensky *et al*. [@partenskyComparisonPhotosyntheticPerformances2018] found that in the low-light conditions found in the OMZ, MED4, SS120 and MIT9313 all became net O~2~ consumers, suggesting that the low light levels cause the respiratory chain to consume more O~2~ than the photosynthetic electron transport chain generates, and c contributing to maintenance of the low O~2~ environment.

>>>>

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: 'Mireille Savoie'
    institute: mta
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    institute: [mta, ubc]
  - name: Julie Nadeau
    institute: mta
  - name: Laurel Genge
    institute: mta
  - name: Sylwia Śliwińska-Wilczewska
    institute: [mta, ug]
  - name: Maximilian Berthold
    institute: mta
  - name: Naaman M. Omar
    institute: mta
  - name: Ondřej Prášil
    institute: [mta, algatech]
    main: false
  - name: Amanda M. Cockshutt
    institute: [mta, sfx]
    main: false
  - name: Douglas A. Campbell
    institute: mta
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Departments of Biology and Chemistry and Biochemistry, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - ubc: 'Department of Biology, University of British Columbia'
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
  - algatech: 'Institute of Microbiology, Center Algatech, Laboratory of Photosynthesis, Novohradska 237, Trebon, CZ 37981, Czech Republic.'
  - ug: 'Institute of Oceanography, University of Gdansk, 46 Piłsudskiego St, Gdynia, Poland'
date: "`r format(Sys.Date())`"
output:
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: FALSE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

Three xenic cultures of *P. marinus* were obtained from Bigelow Labs, NCMA Maine, USA. MED4 (CCMP1986) is from High-Light adapted (HLI) clade; SS120 (CCMP1375) is from Low-Light adapted (LLIII) clade; and MIT9313 (CCMP2773) is from Low-Light adapted (LLIV) clade. Cultures were maintained in incubators set to 22°C with a light/dark cycle of 12 h. The PAR level for maintenance cultures reflected PAR in the source niche of the ecotype; MED4, of 160 µmol photons m^-2^ s^-1^ with illumination from STANDARD Products Inc. Cool White F24T5/41K/8/HO/PS/G5/STD, 24 watts, fluorescent bulbs; SS120 and MIT9313 at 30 µmol photons m^-2^ s^-1^ with illumination from Philips Cool White F14T5/841 Alto, 14 watts, fluorescent bulbs. To maintain active growth all strains were transferred weekly with 1 in 5 dilutions with Pro99 media [@mooreCulturingMarineCyanobacterium2007] prepared with autoclaved artificial seawater [@bergesEvolutionArtificialSeawater2001].

@@

XXX Some text is relevant when referring to original NATLProchlorPhotosynDepthO2 figure, not with this figureXXX
Proteins from 13 annotated strains of *P. marinus* were detected across depths and oxygen concentrations in the ocean proteins data set analyzed. We focused our analysis here on photosynthetic protein complexes, grouped by clade. Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) plots the observation of core photosynthetic complexes (Photosystem II, the Cytochromeb6f complex, Phototsystem I, ATP Synthase and Rubisco) across clades, as a function of depth (a proxy for light intensity) and measured [O~2~]. Photosynthetic complexes from HLI (ex. MED4) were detected throughout the water column, predominantly at high [O~2~]. Complexes from Clade LLIII (ex. SS120) were detected throughout the water column in high [O~2~] and at depth in low [O~2~] samples. Photosystem I and II were only detected at low [O~2~] XXX This is evident when referring to original NATLProchlorPhotosynDepthO2 figure, not with this figureXXX. The exclusive assignment of ATP Synthase peptides to Clade IV (ex. MIT9313) may result from the high conservation of protein sequence for this complex.

@@

XXXRewrite caption if we change figure versionXXX
XXX Figure changed- need to rewrite caption if keeping this figure XXX
```{r CladeProchloroPhotosynthDepthO2, fig.cap = CladeProchloroPhotosynthDepthO2_cap , fig.height = 20, fig.width = 10, echo = FALSE}

CladeProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (clade HLI), SS120 (clade LLIII) and MIT9313 (clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

knitr::include_graphics(file.path(FiguresIn, "CladeProchloroPhotosynthDepthO2.png"))
```

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:AllStrainPurFitsPlots)A). In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)A).  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)A), consistent with Murphy *et al*. [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct fits for different photoperiods refer to figure \@ref(fig:AllStrainPurFitsPlots)A.

@@

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments (Figure \@ref(fig:AllStrainPurFitsPlots)B). Most 4 h photoperiod treatments of SS120 also did not grow under 250 µM O~2~, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments (Figure \@ref(fig:AllStrainPurFitsPlots)B). SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments (Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel PUR for cultures under 25 µM O~2~, showing that SS120 is partially protected from photoinhibition of growth by 25 µM O~2~.  Under 25 µM O~2~, red light again generated more growth of SS120 per unit cumulative diel PUR, than did blue light, again consistent with lower cost of growth through lower photoinactivation under red light (Figure \@ref(fig:BluevsRedPurFitsPlots)B). For distinct fits for different photoperiods refer to figure \@ref(fig:AllStrainPurFitsPlots)B.

@@

The LLIV clade representative, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, lower, flatter response to cumulative diel PUR under 2.5 µM O~2~ (Figure \@ref(fig:AllStrainPurFitsPlots)C). Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands(Figure \@ref(fig:AllStrainPurFitsPlots)C).  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).

Our data again support enhanced growth under conditions of low cumulative diel PUR and 660 nm (red) spectral bandwidth, consistent with Murphy *et al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost of growth, due to decreased photoinactivation of PSII under red, compared to blue, wavebands (Figure \@ref(fig:BluevsRedPurFitsPlots)C).  Interestingly, this protective effect of red light disappears for MIT9313 growing under 2.5 µM O~2~, possibly because photoinactivation is strongly suppressed under this low [O~2~]. For distinct fits for different photoperiod fits refer to figure \@ref(fig:AllStrainPurFitsPlots)C.

@@

\newpage
XXXXEdit text depending upon which ocean protein figures we useXXXXX
Ocean detections of proteins mediating protein metabolism support this interpretation of distinct FtsH function across clades of *P. marinus*. Ribosome proteins from clade HLI MED4, clade LLIII SS120 and clade LLIV MIT9313 show generally similar patterns vs. [O~2~] and depth, a proxy for peak PAR.  FtsH3, inferred to mediate PSI assembly, likewise shows a similar pattern between MED4 and MIT9313.  But only MED4 shows detected presence of the FtsH1 & FtsH2 isoforms inferred to mediate PSII repair, and then only in near-surface samples subject to higher light.  Furthermore, even though MIT9313 grows (Figure \@ref(fig:MIT9313GAM)), and is detected in the ocean at low [O~2~] (Figure \@ref(fig:CladeProchloroPhotosynthDepthO2)), no FtsH from MIT9313 is detected at low [O~2~], suggesting limited requirement for protein turnover under low [O~2~].

@@

*Prochlorococcus marinus* MED4 has a physiological requirement for more than 4 h of light per day, and thus this strain will not exploit habitats with short photoperiods, typical of winter or light attenuated depths. MED4 is also excluded from the lowest oxygen habitats, equivalent to our 2.5 µM O~2~ experimental conditions. MED4 can, however, grow under OMZ regions with slightly higher [O~2~], as demonstrated by the growth under our 25 µM O~2~ experiments. Genomic analyses (Figure \@ref(fig:ProchlorococcusEnzymeKms))  and previous transcriptional analyses [@bergResponsesPsbAHli2011]  suggest MED4 is excluded from growth below ~ 25 µM O~2~ because it relies upon a ubiquinol oxidase, non-electrogenic, to maintain oxidation/reduction balance in the intersystem electron transport chain.  On the other hand, MED4 shows inducible expression of FtsH isoforms [@bonisteelStrainSpecificDifferences2018], to counter photoinactivation of PSII under higher PAR and [O~2~] environments. Photoinactivation, does, however, impose an increased cost of growth upon MED4, since growth under red light, to lower the rate of photoinactivation of PSII [@murphyPhotoinactivationPhotosystemII2017], allows MED4 to achieve faster growth per absorbed photon than growth under red light. TARA Oceans
Project data [@delmontLinkingPangenomesMetagenomes2018] indeed reported presence of *P. marinus* MED4-like genomes at depths ranging from 5 m to 90 m, representing high to low blue light levels, in the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018 did not analyze data from depths beyond the subsurface chlorophyll maximum layer, the layer in the water column where the chlorophyll *a* concentration peaks, nor did they report the [O~2~] at those depths. Our growth findings are also consistent with Figure\@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII proteins annotated as MED4, clade HLI, at depths up to 200 meters with O~2~ of 15 µM.

*Prochlorococcus marinus* SS120, a LLIII clade representative, showed an interactive inhibition of growth by oxygen and cumulative diel PUR, with a higher tolerance for higher cumulative diel PUR under 25 µM O~2~, compared to 250 µM O~2~. Thus, SS120 can exploit habitats with O~2~ levels spanning 25 µM O~2~ to 250 µM O~2~, including high PAR environments within OMZ.  SS120 is likely excluded from higher [O~2~] and higher PAR because of genomic limitations on capacity for DNA repair (Figure \@ref(fig:DNARepairFig)), and possibly by limited capacity for synthesis of reactive oxygen quenchers (Figure  \@ref(fig:ProchlorococcusEnzymeKms)).  Our growth results are supported by @lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length polymorphism analyses, at depths above 40 m, where light levels are
higher, within OMZ, and by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSII protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in an OMZ of the tropical North Pacific Ocean. SS120, grew under photoperiods longer than 4 h and showed increasing growth rate with increasing photoperiods. Therefore, SS120 can potentially exploit the longer summer photoperiods in temperate zones but again cannot sustain growth in winter through its requirement for more than 4 h of light per day, even if water temperature warms into the clade LLIII tolerance range.

@@

*Prochlorococcus marinus* MIT9313, a LLIV clade representative, shows the potential to inhabit future warmer temperate zones year-round, as it grows under a 4 h photoperiod, expected in winter or at light-attenuated depths. MIT9313 demonstrates an unexpected tolerance to higher light levels and cumulative diel PUR,  but only under low oxygen conditions of 25 µM and 2.5 µM (Figure \@ref(fig:MIT9313PAR)). The ability to utilize higher PAR and higher cumulative diel PUR under 25 µM or 2.5 µM O~2~ enables MIT9313 to grow in OMZ, even at depths closer to the surface. MIT9313 carries a gene encoding (S)-2-hydroxy-acid oxidase [@cunaneCrystalStructureAnalysis2005],  with a K~M~ for [O~2~] of ~ 250 µM (Figure \@ref(fig:ProchlorococcusEnzymeKms)), which produces H~2~O~2~.  Growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffers photoinhibition, resulting from the inactivation of PSII caused by the production of H~2~O~2~. This photoinhibition is compounded by the limited inducible repair for PSII, due to the absence of FtsH 2 and 3  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  MIT9313 shows remarkable ability to thrive under very low [O~2~], potentially allowing it to expand into broader ecological niches. These results are supported by Figure \@ref(fig:CladeProchloroPhotosynthDepthO2) showing PSI protein subunits annotated as derived from MIT9313 detected at depths \> 120 m, along with PSII subunits at depths from 50 m to 200 m in regions where O~2~ was 15 µM.

In future work we aim to test whether MIT9313 is not “using” photosynthesis in low O~2~ environments to drive CO~2~ fixation, so much as to generate oxygen locally to support  respiration. The deep water environments typical for MIT 9313 are relatively nutrient rich and  *Prochlorococcus* take up and metabolize various sugars [@gomez-baenaGlucoseUptakeIts2008; @munoz-marinGlucoseUptakeProchlorococcus2017; @munoz-marinProchlorococcusCanUse2013] and amino acids [@zubkovDepthRelatedAmino2004]. They may be using photosynthesis to generate O~2~ to act as an electron sink for respiration, to generate using ATP for maintenance and to take up nutrients from the surroundings.
Partensky *et al*. [@partenskyComparisonPhotosyntheticPerformances2018] found that in the low-light conditions found in the OMZ, MED4, SS120 and MIT9313 all became net O~2~ consumers, suggesting that the low light levels cause the respiratory chain to consume more O~2~ than the photosynthetic electron transport chain generates, and c contributing to maintenance of the low O~2~ environment.
